Co-moving stars in \textsl{Gaia DR1}:\\ An abundance of very wide separation co-moving pairs

### 2016-12-05 paper/ms.tex
final edits
- thickened lines in histogram figures
- lowererd alpha for the initial sample plot

\begin{abstract}
The primary sample of the \gaia\ Data Release 1 is the
\textsl{Tycho-Gaia Astrometric Solution} (\tgas): $\approx$ 2 million Tycho-2
sources with improved parallaxes and proper motions relative to the initial
catalog.
This increased astrometric precision presents an opportunity to find new binary
stars and moving groups.
We search for high-confidence co-moving pairs of stars in \tgas\ by identifying
pairs of stars consistent with having the same 3D velocity using a marginalized
likelihood ratio test to discriminate candidate co-moving pairs from the field
population.
Although we perform some visualizations using (bias-corrected) inverse-parallax
as a point-estimate of distance, the likelihood ratio is computed with a
probabilistic model that includes the covariances of parallax and proper motions
and marginalizes the (unknown) true distances and 3D velocities of the stars.
We find 13,085 co-moving star pairs among 10,606 unique stars with separations
as large as 10 pc (our search limit).
Some of these pairs form larger groups through mutual co-moving neighbors:
many of these pair networks correspond to known open clusters and OB
associations, but we also report the discovery of several new co-moving groups.
Most surprisingly, we find a large number of very wide ($>1$ pc) separation
co-moving star pairs, the number of which increases with increasing separation
and cannot be explained purely by false-positive contamination.
Our key result is a catalog of high-confidence co-moving pairs of stars in
\tgas.
We discuss the utility of this catalog for making dynamical inferences about the
Galaxy, testing stellar atmosphere models, and validating chemical abundance
measurements.
\end{abstract}

@@

At larger separations ($\gtrsim 1$~pc), co-moving stars are likely members of
(potentially dissolving) moving groups, associations, and star clusters or
disrupted wide binaries.
The origin of moving groups is still under active debate (e.g., \citealt{Bovy:2010aa}:
are they remnants of a coeval
star formation event with similar chemical composition? Or are they formed by
dynamical effects of non-axisymmetric features of the Galaxy such as spirals
and bars? With the recent advances in measuring chemical abundances of
a large volume of stars using high and low resolution spectroscopy
(e.g., \citealt{Steinmetz:2006aa,Majewski:2015aa,Gilmore:2012aa} to name a few),
we can now start to explore these questions
with unprecedented statistics and in unexplored detail.
The dynamics of cluster dissolution provides important clues to understanding
the star formation history and the dynamical evolution of the Milky Way.
In the halo, we know of more than 20 disrupting globular clusters and dwarf galaxies
(``stellar streams''; see, e.g., \citealt{Grillmair:2016} for a summary of known streams).
These tidal streams are modeled to infer the parameters of the Galactic
potential (e.g., \citealt{Kupper:2015}).
Similar processes are at work with star clusters in the disk.
However, the dynamical time is much shorter, and the dynamics will be much more
complex because of the existence of other perturbers in the disk.

>>>>

\begin{abstract}
The primary sample of the \gaia\ Data Release 1 is the
\textsl{Tycho-Gaia Astrometric Solution} (\tgas): $\approx$ 2 million Tycho-2
sources with improved parallaxes and proper motions relative to the initial
catalog.
This increased astrometric precision presents an opportunity to find new binary
stars and moving groups.
We search for high-confidence co-moving pairs of stars in \tgas\ by identifying
pairs of stars consistent with having the same 3D velocity using a marginalized
likelihood ratio test to discriminate candidate co-moving pairs from the field
population.
Although we perform some visualizations using (bias-corrected) inverse-parallax
as a point-estimate of distance, the likelihood ratio is computed with a
probabilistic model that includes the covariances of parallax and proper motions,
and marginalizes the (unknown) true distances and 3D velocities of the stars.
We find 13,085 co-moving star pairs among 10,606 unique stars with separations
as large as 10 pc (our search limit).
Some of these pairs form larger groups through mutual co-moving neighbors:
many of these pair networks correspond to known open clusters and OB
associations, but we also report the discovery of several new co-moving groups.
Most surprisingly, we find a large number of very wide ($>1$ pc) separation
co-moving star pairs, the number of which increases with increasing separation
and cannot be explained purely by false-positive contamination.
Our key result is a catalog of high-confidence co-moving pairs of stars in
\tgas.
We discuss the utility of this catalog for making dynamical inferences about the
Galaxy, testing stellar atmosphere models, and validating chemical abundance
measurements.
\end{abstract}

@@

At larger separations ($\gtrsim 1$~pc), co-moving stars are likely members of
(potentially dissolving) moving groups, associations, and star clusters or
disrupted wide binaries.
The origin of moving groups is still under active debate (e.g., \citealt{Bovy:2010aa}):
are they remnants of a coeval
star formation event with similar chemical composition? Or are they formed by
dynamical effects of non-axisymmetric features of the Galaxy such as spirals
and bars? With the recent advances in measuring chemical abundances of
a large volume of stars using high and low resolution spectroscopy
(e.g., \citealt{Steinmetz:2006aa,Majewski:2015aa,Gilmore:2012aa} to name a few),
we can now start to explore these questions
with unprecedented statistics and in unexplored detail.
The dynamics of cluster dissolution provides important clues to understanding
the star formation history and the dynamical evolution of the Milky Way.
In the halo, we know of more than 20 disrupting globular clusters and dwarf galaxies
(``stellar streams''; see, e.g., \citealt{Grillmair:2016} for a summary of known streams).
These tidal streams are modeled to infer the parameters of the Galactic
potential (e.g., \citealt{Kupper:2015}).
Similar processes are at work with star clusters in the disk.
However, the dynamical time is much shorter, and the dynamics will be much more
complex because of the existence of other perturbers in the disk.
